---
title: "ECON 6760 - Assignment 1"
author: "Joe Martin"
date: "1/17/2022"
output: word_document
---

```{r setup, include=FALSE}
pacman::p_load(pacman, tidyverse, tidymodels, lubridate, here, tidytuesdayR)

#Get the Data

# Read in with tidytuesdayR package 
# Install from CRAN via: install.packages("tidytuesdayR")
# This loads the readme and all the datasets for the week of interest

# Either ISO-8601 date or year/week works!

#tuesdata <- tidytuesdayR::tt_load('2022-01-11')
#tuesdata <- tidytuesdayR::tt_load(2022, week = 2)

#colony <- tuesdata$colony

# Or read in the data manually

colony <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/colony.csv')
stressor <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/stressor.csv')

months <- data.frame(month_name = c("January","April","July", "October"), month_num = c(1,2,3,4))

# Create a new variable to display "Months" as "QUARTERS" 
colony <- colony %>% 
  separate(col = months,into = c("first_m","last_m"), sep = "-")
colony$ym <- paste(colony$year,colony$first_m, sep = "-")
colony$ym <- ym(colony$ym)
colony$quarter <- quarter(colony$ym)
```

Explore the data. For reference, this is the meaning for variables:

colony_n: Number of colonies
colony_max: Maximum colonies
colony_lost: Colonies lost
colony_lost_pct: Percent of total colonies lost
colony_added: Colonies added
colony_reno: Colonies renovated
colony_reno_pct: Percent of colonies renovated
```{r}
# How many colonies were lost over time in Georgia
ts_colony_ny <- colony %>%
  filter(state == "New York") %>%
  select(ym,colony_n)

ts_colony_ga <- colony %>%
  filter(state == "Georgia") %>%
  select(ym,colony_n)

ts_colony_ny <- ts(ts_colony_ny)
ts_colony_ga <- ts(ts_colony_ga)

plot(ts_colony_ga)
plot(ts_colony_ny)
```
Although this is not economic data, this dataset is recorded quarterly. This is interesting because it is possible to see seasonal population trends. For example, filtering for colony numbers in Georgia, it is less clear how colony numbers are affected by weather, likely due to milder autumns and winters. However, when filtering for a state like New York where there are more dramatic changes in seasonal temperatures, it is clear that populations increase and decrease depending on the season.

```{r}
# Is this enough? Should I create a larger dataset for the whole US and see write about those trends?
```

```{r}

colony %>%
  filter(state == "Georgia") %>%
  ggplot(aes(x=ym, y = colony_lost))+
  geom_smooth()
```

```{r}
# Try using a natural log scale 
colony %>%
  filter(state == "Georgia") %>%
  ggplot(aes(x = ym, y = log10(colony_lost)))+
  geom_smooth()
```

```{r, include=FALSE}
# Download the monthly data series CPIAUCSL, the seasonally adjusted consumer price index for all urban consumers in the US, from the St. Louis Fed Data Base (FRED).
cpia <- read_csv("~/MSBA/Spring 2022/ECON 6760/econ-6760/CPIAUCSL.csv")
cpia$month <- lubridate::month(cpia$DATE)
cpia$year <- lubridate::year(cpia$DATE)
cpia$quarter <- lubridate::quarter(cpia$DATE)
cpia$ln_CPIA <- log10(cpia$CPIAUCSL) #Transform to natural log scale
#cpia$first_diff <- diff(cpia$ln_CPIA)
```

```{r}
# Show plot as it appears on FRED website (https://fred.stlouisfed.org/series/CPIAUCSL)
cpia %>%
  ggplot(aes(x = year, y = CPIAUCSL))+
  geom_smooth(se = FALSE)+
  scale_x_continuous(breaks = seq(1950,2021, by = 5))
  
```

```{r}
# natural log scale
cpia %>%
  ggplot(aes(x = year, y = ln_CPIA))+
  geom_smooth(se = FALSE)+
  scale_x_continuous(breaks = seq(1950,2021, by = 5))
```

```{r}
ts_cpia <- cpia %>%
  select(DATE,CPIAUCSL)

ts_cpia <- ts(ts_cpia)

plot(diff(log10(ts_cpia)))

plot(ts_cpia)
```